Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2021-10-12 doc/AMS-15/AMS-15_poster.Rmd
Put together figures for AMS-15 poster

---
title: "HGIS Carbonate Results"
subtitle: "NAPB 2021"
author: 
  - "Brett Longworth"
date: '2021-10-08'
---

@@

```{r}
library(tidyverse)
library(amstools)
library(hgis)
library(here)
library(gt)
library(Bchron)
library(gghighlight)

@@

## Method

@@

## Results: Agreement with Consensus

@@

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))
```

>>>>

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
- Brett Longworth
- Josh Burton
- Simon Pendleton
- Josh Hlavenka
- Sydney Moser
- Mark Roberts
- Mark Kurz
output:
  html_document: default
  word_document: default
  pdf_document: default
---

@@

```{r}
library(tidyverse)
library(amstools)
library(hgis)
library(here)
library(gt)
library(Bchron)
library(gghighlight)
library(patchwork)

@@

Text and figures for AMS-15 poster.

# Abstract

We have developed a method for measurement of radiocarbon in carbonate minerals
as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
uses helium as a carrier gas to displace CO~2~ from sample vials to an open
split, where a glass capillary samples the mixture for delivery directly to the
HGIS. This method skips the gas transfer and quantification steps used in a
closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
mg carbonate can currently be measured. Results from measurements of consensus
standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
from a marine core (40-115 pMC) show that the method agrees well with
traditional AMS measurement of the same samples as graphite, and that the method
error is less than 1.5 pMC. We discuss advantages and disadvantages of
continuous flow sample introduction, and the effect of reduced precision on
calibrated age-depth models produced using gas-source data.

# Introduction

# Method

@@

# Results

Data used for current, blanks, consensus

@@

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))
```
## Initial measurements

### Flow vs current

```{r cur-time-os, fig.height=7, dev='svg'}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
  filter(Pos == 5) %>% 
  mutate(time = cum_acqtime/60,
         co2flow = concCO2(time, r = 244) * 30) #30ul/min

# flow vs time subplot
flow_time <- data.frame(x = 0:125) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2"), size = 1.5) +
  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium"), size = 1.5) +
  scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Gas flows to source",
       subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
       y = "Gas flow (μL/min)") +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.87, 0.65),
        legend.background = element_rect(fill = "white", color = "black"))

# current vs time subplot
cur_time <- ggplot(data, aes(time, he12C)) +
  geom_smooth(span = .4, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  xlim(0, 125) +
  labs(title = "Ion current",
       x = "Time (min)",
       y = "12C current (μA)") +
  theme_classic()

flow_time / cur_time
```

```{r cur-flow-os, fig.height=6, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  labs(title = "Vial dilution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic()
```

compare to first tests

Note that our flat-topped current curve allows variable flow of the vial system. MICADAS has a more pronounced peak so this may not work as well. (Check papers for peak shape)

Also note current dependent ratio (blank) as a complicating factor

Lead into blank on next slide

### Blanks and blank correction

# Current independent blank

If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.

#### Linear fit used to correct:

$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$

#### Mean Fm of blanks: `r cons_df %>% filter(fm_consensus < 0.1 & he12C > 2) %>% pull(norm_ratio) %>% mean() %>% format(digits = 3)`

```{r blanks, fig.height=5, dev='svg'}
cons_df %>% 
  filter(fm_consensus < 0.1) %>% 
  ggplot(aes(he12C, norm_ratio)) +
  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio  + sig_norm_ratio)) +
  geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C  + sig_he12C)) +
  geom_point(aes(color = wheel), size = 3) +
  #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
  labs(title = "Blanks",
       x = "12C Current (μA)",
       y = "Fraction modern") +
  theme_classic() +
  theme(legend.position = c(0.82, 0.76),
        legend.background = element_rect(fill = "white", color = "black")) 
```

## Results: Consensus standards
